\documentclass[12pt]{article}
\usepackage{guide_common-preamble} % guide_common-preamble.sty

% Author setting
\usepackage{authblk}

\title{\Large\textbf{Candidates in American General Elections}}

\author[1]{\normalsize Jeremiah Cha}
\author[2]{\normalsize Shiro Kuriwaki}
\author[1]{\normalsize James M. Snyder, Jr.}
\affil[1]{\footnotesize Department of Government, Harvard University.}
\affil[2]{\footnotesize Department of Political Science and Institution of Social and Policy Studies, Yale University}
% \affil[*]{\url{https://github.com/kuriwaki/cces_candidates} or email \href{mailto:shirokuriwaki@gmail.com}{\texttt{shirokuriwaki@gmail.com}}}


\date{\vspace{-0.2in} \normalsize Version 3.0 (October 2024)}


\begin{document}
\maketitle

\vspace{-0.2in}

\begin{center}
\begin{minipage}{0.75\linewidth}
\small

\emph{Please cite as: }

Cha, Jeremiah, Shiro Kuriwaki, and James M. Snyder, Jr., 2024, ``Candidates in American General Elections'', Harvard Dataverse. \textsc{doi}: \href{https://doi.org/10.7910/DVN/DGDRDT}{\texttt{10.7910/DVN/DGDRDT}}.
\end{minipage}
\end{center}
\smallskip

Between 2006 and 2022, there were over 4,000 races for Congress and Governor in the general election, in which \input{Ns/all_cands}\unskip \ Democratic and Republican candidates ran for office.
While many existing efforts compile these election results, few include standardized candidate names across years and offices, candidate incumbency status, and elections for different offices and years combined into a single file.

We provide a candidate-level dataset that is comprehensive for Presidential, Congressional and gubernatorial contests between 2006 and 2022 with standardized names for each unique candidate. We also record party, incumbency, vote totals, and election results. For example:

\begin{table}[!h]
\centering
\footnotesize
\input{Tables/cands/example_candidates}
\end{table}

\newpage

\section{Data Sources and Coverage}

Most of the candidate-level data -- names and incumbency status -- is a subset of the data collected and manually coded by James M. Snyder, Jr. Vote counts and candidate listings of Congress are entered from the Official Election Statistics compiled by the House Clerk (\url{https://history.house.gov/Institution/Election-Statistics/Election-Statistics/}).

Election results for the state office of the Governor are collected from statements of votes by the Secretary of State of each state. We include the odd-year elections in Virginia, New Jersey, Kentucky, Mississippi, and Louisiana.
Recall elections (Wisconsin 2012, California 2021) are not included.

Presidential vote counts are taken from the data compiled by the MIT Election Data Science Lab (\url{https://doi.org/10.7910/DVN/42MVDX}), which in turn mostly sources from the Official Election Statistics by the House Clerk.

Incumbency is determined by a programmatic count of whether the name is repeated across years, and by manual inspection of the biography of each candidate.

This dataset excludes candidates who receive less than ten votes in their general election and those who identify Blank or Miscellaneous as their party.
We also exclude overvotes, undervotes, and invalid votes.
Although these removals make the \texttt{totalvotes} be a few votes less than the actual turnout, we made this choice uniformly because states vary in their reporting protocol for such votes. That is, given that some states report overvotes while others exclude them entirely in their official counts, it is more internally consistent to exclude them for all states.

Our code for recoding and editing the data is open-source in a Github repository (\url{https://github.com/kuriwaki/cces_candidates}).


\section{Unique Identifiers}

To identify a single election, group by \texttt{year}, \texttt{state}, \texttt{office}, \texttt{dist}, and \texttt{type} (General vs. Special).  To isolate the set of candidates who won a seat in an election cycle, group by subset by \texttt{year}, \texttt{state}, \texttt{office}, and \texttt{type}, then subset by \texttt{won}.

To identify a single candidate across office and time, group by \texttt{state} and  \texttt{name\_snyder}.
As we discuss in detail in this documentation, the candidate names are standardized (by Snyder's coding) so that the same candidate has the same spelling across years and office.
In very rare instances, a candidate for a major party runs for office two different states. For example, Mitt Romney ran for Governor of Massachusetts, President, and U.S. Senator from Utah. For these candidates, group by \texttt{name\_snyder}.
% \clearpage
\section{Variable Descriptions and Tabulations}



\subsection*{\texttt{office}}

The office the candidate is running for. Following the Snyder data, we use \texttt{H} for the US House of Representatives, \texttt{S} for the US Senate, and \texttt{G} for Governor.


    \begin{tabularx}{\linewidth}{C}
    \input{Tables/cands/office}
    \end{tabularx}

\subsection*{\texttt{dist}}

The district the candidate is running for. U.S. House races in at-large districts are numbered 1 instead of 0 or ``-AL''.

 For U.S. Senate candidates, the district variable denotes the class of the particular Senate seat (e.g. Dianne Feinstein has a \texttt{dist} value of 1 since her seat is part of Senate Class I).

\begin{table}[!h]
\centering
\footnotesize
\input{Tables/cands/dist_sum}
\end{table}





\subsection*{\texttt{type}}

The type of election the candidate is running in. We use \texttt{G} to denote general election candidates and \texttt{S} to denote special election candidates. Candidates are said to run in special elections if they are running to serve out the term of another legislator.
\begin{table}[!h]
\begin{tabularx}{\linewidth}{CCC}
    \textbf{Senate} & \textbf{House} & \textbf{Governor}\\
    \input{Tables/cands/type_S} & \input{Tables/cands/type_H} & \input{Tables/cands/type_G}
\end{tabularx}
\end{table}

\FloatBarrier

\subsection*{\texttt{runoff}}

The candidate's election round. In some states (e.g. Louisiana), if no candidate receives over 50 percent of the vote, the top two candidates must run in a runoff election again to determine the winner.
In such a case, \emph{only the deciding contest is included}. For example, in the 2020 Georgia Senate elections, both the general and special election went to a runoff because no candidate got over 50 percent of the vote in either race. Therefore, we exclude the November election results and include the January runoff.
A runoff variable with a value of 1 denotes that the election the candidate is running in is a runoff election, while a 0 denotes that the election is not. A missing value is given for the vast majority of states in which no such runoff option is available, and a candidate can win by plurality.
\begin{table}[!h]
\begin{tabularx}{\linewidth}{CCC}
    \textbf{Senate} & \textbf{House} & \textbf{Governor}\\
    \input{Tables/cands/runoff_S} & \input{Tables/cands/runoff_H} & \input{Tables/cands/runoff_G}
\end{tabularx}
\end{table}


\FloatBarrier

\subsection*{\texttt{name\_snyder}}


Standardized candidate name following Snyder coding rules. The
specifics of these coding rules are important for overtime tracking because
\textbf{name and state uniquely identify a unique person} across all his
Congressional candidate data (at least 1950 to the present). The syntax
is

\texttt{{[}Last\ name{]},\ {[}First\ Name{]}\ {[}Middle\ name{]}\ ({[}Nickname{]}),\ {[}JR./SR./I/II/III{]}}

For example:

\begin{itemize}
\tightlist
\item
  \texttt{SEWELL,\ TERRYCINA\ ANDREA\ (TERRI)}: commonly known as Terri
  Sewell (AL)
\item
  \texttt{GRASSLEY,\ CHARLES\ ERNEST\ (CHUCK)}: commonly known as Chuck
  Grassley (IA)
\item
  \texttt{CORNYN,\ JOHN,\ III}: commonly known as John Cornyn (TX)
\item
  \texttt{KENNEDY,\ JOSEPH\ P.\ (JOE),\ III}: commonly known as Joe
  Kennedy (MA)
\item
  \texttt{KELLY,\ GEORGE\ J.\ (MIKE),\ JR.}: commonly known as Mike
  Kelly (PA). Note the use of periods for abbreviation and comma before
  the JR.
\item
  \texttt{WASSERMAN\ SCHULTZ,\ DEBBIE}: note the last name is not
  hyphenated and is two words
\item
  \texttt{JACKSON\ LEE,\ SHEILA}: note that the last name is not Lee, so
  we use this instead of ``Lee, Sheila Jackson''
\end{itemize}

Note that these names are \emph{different} from other names (1)
on the ballot, (2) on the Clerk of the House report, or the (3) FEC. Any
of these sources can disagree on the formatting of the names or whether
to abbreviate a middle name. The same candidate in the same election year may
even have different formats in each of these sources.

The Snyder formalization solves some shortcomings in the official data
above. Candidates may report different style names on the ballot vs.~the
FEC or they may have multiple FEC IDs at the same time. More commonly, a
House representative who then runs for higher office like the Senate will
often get a separate FEC ID, even though they are the same person. In
theses cases, the Snyder formalization of identifying people by their full
name and state is the most reliable.

Therefore, it is important in our formalization to spell out nicknames
and distinguish between hyphenated last names to the extent that it
uniquely identifies a person within a state. The same person's name
should be formatted exactly the same way, and conversely, two different
people with near-identical last names should somehow be given
{different} names. For example, if a John Smith ran for Congress
in 1980 in Ohio and then a \emph{separate, unrelated} John Smith ran
again in Ohio in 2020. To distinguish the two different people, we
would need to find {something} like a middle name or suffix to
distinguish the name in characters.

One shortcoming of the Snyder formalization is that it may be
{too} rigid. Whether or not something is abbreviated can be
determined on a coding rule decided early on. If a person's name
changes, e.g.~through marriage or divorce, the old name must be kept.
For example:

\begin{itemize}
\tightlist
\item
  \texttt{BIDEN,\ JOSEPH\ R.,\ JR.} has his middle name abbreviated,
  even though it is known (Robinette).
\item
  \texttt{MACK,\ MARY\ BONO} (R-CA, 1998-2012) married to Connie Mack IV
  (R-FL, 2001-2003) during 2007-2013. They separated in 2013 and she now
  goes by Mary Bono, but in the data she will be coded by the above
  name.
\end{itemize}


\newpage

\FloatBarrier


\subsection*{\texttt{party}}

The short party affiliation of the candidate. We use the ``short'' or colloquial party name. For example, the Democrat-Farmer-Labor Party in Minnesota is given a \texttt{D} instead of \texttt{DFL}. Candidates who ran on third party tickets in Connecticut and New York are simply given the major party name.
\begin{table}[!h]
\begin{adjustwidth}{-4em}{-4em}
\centering
\begin{tabularx}{\linewidth}{CCC}
\textbf{Senate} & \textbf{House} & \textbf{Governor}\\
  \input{Tables/cands/party_S} & \input{Tables/cands/party_H} &\input{Tables/cands/party_G}
\end{tabularx}
\end{adjustwidth}
\end{table}


\FloatBarrier

\subsection*{\texttt{party\_formal}}
The formal name of the candidate's party affiliation, for example DFL in Minnesota. Third party candidate names follow what is given in the House Clerk document or state SOS reports (for Governors), with some minor abbreviations and standardizations.

\twocolumn
\begin{adjustwidth}{-2em}{-2em}
\centering
{\scriptsize
\input{Tables/cands/party-formal_by-office}
}
\end{adjustwidth}

\onecolumn



\subsection*{\texttt{inc}}

The incumbency status of the candidate. A 0 means the candidate is not an incumbent, a 1 means the candidate is an incumbent, a 2 means the candidate is an incumbent that was elected in a special election, and 3 indicates a U.S. Senator who was appointed by the Governor.
\begin{itemize}
\item District numberings do not matter for determining incumbency. For example, Conor Lamb who won a special election to PA-18 in March 2018 is still an incumbent when he ran in November 2018 for the re-drawn PA-17.
\item This may differ from \textbf{the CES incumbency variable} in some redistricting cases when two incumbents were forced to run against each other. For example, in 2012 Betty Sutton (OH-13) and Jim Renacci (OH-16), both House incumbents, were forced to run in OH-16. While the CES incumbency variable (HouseCandIncumbent) lists only Sutton as the incumbent, this dataset will list both Sutton and Renacci as incumbents.
\item If a special election and general election are held on the same day (e.g., NY-29's House election in 2010), the special election results on incumbency (inc = 2) do not carry over to the general election incumbency status.
\end{itemize}

    \begin{tabularx}{\linewidth}{CCC}
    \textbf{Senate} & \textbf{House} & \textbf{Governor}\\
    \input{Tables/cands/inc_S} & \input{Tables/cands/inc_H} & \input{Tables/cands/inc_G}
    \end{tabularx}


\FloatBarrier


\subsection*{\texttt{nextup}}

The year that an incumbent candidate is up for re-election. This variable can be useful in determining tenure in office for special election candidates, who may not have served a full term. See \texttt{type} for designation of special election candidates.

\begin{tabularx}{\linewidth}{C}
\textbf{Senate}\\
\input{Tables/cands/nextup_S}
\end{tabularx}

\begin{tabularx}{\linewidth}{C}
\textbf{House}\\
\input{Tables/cands/nextup_H}
\end{tabularx}

\begin{tabularx}{\linewidth}{C}
\textbf{Governor}\\
(Varies, with NJ, VA, MS, LA, KY having odd year elections and NH, VT having two year terms)
\end{tabularx}

\FloatBarrier

\subsection*{\texttt{candidatevotes}}

The number of total votes the candidate received.
\begin{itemize}
\item For candidates running on multiple party tickets, this will be the \emph{sum} of all of their votes.  For example, in 2016, Rep. Rosa L. DeLauro (CT-03) ran as a Democrat and also ran as a Working Families Party candidate. She won 192,274 votes in the former and 21,298 votes in the latter, so her \texttt{candidatevotes} is the total, 213,572.
\item  Florida and Oklahoma do not report the vote count for a House candidate if she won unopposed. In these cases, we have the vote count as \texttt{NA} but have the candidate winning (\texttt{won == 1}).
\end{itemize}

Therefore, the data for such corner cases look like:

\begin{table}[!h]
\centering
\footnotesize
\input{Tables/cands/example_candidatevotes}
\end{table}


The total distribution for a candidate's votes looks as follows:

\begin{table}[!h]
\centering
\footnotesize
\input{Tables/cands/candidatevotes_sum}
\end{table}


\FloatBarrier

\subsection*{\texttt{totalvotes}}

Total votes for all candidates in each election. This is the sum of \texttt{candidatevotes}. It does not include undervotes, overvotes, scattered votes or votes for candidates with less than 10 votes.  Also note that only deciding elections are included in the data (see \texttt{runoff}). Therefore for runoff elections, \texttt{totalvotes} reflects the total turnout in the second round, which often has lower turnout than the November election.



\subsection*{\texttt{won}}
Candidate won the general election. A 1 means the candidate won the general election, and a 0 means the candidate lost the general election.

Presidential candidates win electoral college votes at the state level, but only win office with a majority of the college votes. Because of this ambiguity, we have left this variable missing for Presidential candidates.

    \begin{tabularx}{\linewidth}{CCC}
    \textbf{Senate} & \textbf{House} & \textbf{Governor}\\
    \input{Tables/cands/won_S} & \input{Tables/cands/won_H} & \input{Tables/cands/won_G}
    \end{tabularx}

The above table has more than 435 candidates winning in the House each year because they count special elections held that year too. Here are the counts for only general elections. For 2018, the House winners are not 435 because we coded the results in NC-09 to be missing, due to a voter fraud case where the election was not certified.

\begin{tabularx}{\linewidth}{CCC}
    \textbf{Senate} & \textbf{House} & \textbf{Governor}\\
    \input{Tables/cands/won_S_gen} & \input{Tables/cands/won_H_gen} & \input{Tables/cands/won_G_gen}
    \end{tabularx}


The set of tables below show the \emph{party} of the candidate that \emph{won}.

    \begin{tabularx}{\linewidth}{CCC}
    \textbf{Senate} & \textbf{House} & \textbf{Governor}\\
    \input{Tables/cands/party-won_S} & \input{Tables/cands/party-won_H} & \input{Tables/cands/party-won_G}
    \end{tabularx}

The NC-09 House race in 2018 is listed as having no winner because it was later overturned. A special election was held in 2019.

% \item \texttt{data\_note}: 145 candidates were not included in the election dataset we used to collect the candidate-level variables. We still include these candidates in the supplemental data in case researchers are interested in specific districts or races. The reason for these non-merges can be broken down into four categories, listed below.
% 		\begin{enumerate}
% 			\item Incorrect Election: Candidates with this code  did not run in the respondent's district or state. This likely occurs because the respondent's information was incorrectly entered.
% 			\item Not on General Election Ballot: Candidates with this code either withdrew from the race, were disqualified, were write-in candidates, did not make it to the included runoff election, or did not receive enough votes to appear on the ballot.
% 			\item Unopposed in Oklahoma: In Oklahoma, unopposed candidates do not appear on the general election ballot.
% 			\item Missing from election data: Candidates with this code were not included in the data we used to merge in the election information. Oftentimes, this category is made up of candidates in races with jungle primaries, such as Louisiana. This also includes all candidates from Washington D.C..
% 		\end{enumerate}




\section{Related Work and Extensions}

The MIT Election Data Science Lab (\url{https://dataverse.harvard.edu/dataverse/medsl\_election\_returns}) produces similar election results. Their data goes back as far as 1976. However, the data does not include incumbency status, and the names are taken as officially reported rather than standardized to match across years.


% \section{Extensions: Race, Gender, and Primary Elections}

Our dataset does not contain information about the gender or race of the candidates. We hope to work collect that data and combine other sources of data in the future.  In the meantime there are several related data sources users can rely on. First, Numerous years in the CES have data available on candidate race and gender for interested researchers. Please see the table below for further information regarding the availability of such information by year and where it can be located.

\begin{table}[H]
	\footnotesize
	\centering
	\caption{The Availability of Candidate Race and Gender Data by Year}
\begin{tabularx}{0.7\linewidth}{lXX}
	\toprule
	CES   & Candidate Race &Candidate Gender \\
	\midrule
	2006 &  &\\
	2008 & Variables for H, S, G & Variables for H, S, G  \\
	2010 & Supplemental Data for H$^1$  &  Variables for H, S, G\\
	2012 & Supplemental Data for H$^2$ & Variables for H and S \\
	2014 & Supplemental Data for H$^3$ &\\% Variables for H and S \\[2ex]
	2016 & Supplemental Data for H, S$^4$  & \\%Variables for current H and S  \\[2ex]
	2018 &  & Variables for current H post \\
	\bottomrule
\end{tabularx}
% \caption*{\footnotesize Note that the 2013 common content also includes the gender of House members.}
\bigskip

%\skcomment{I commented out "for current.." because that's not what the column says. I also deleted the odd year rows and used booktabs for table formatting}

\footnotesize
\singlespacing
		1. \url{https://doi.org/10.7910/DVN/KC9EQR} \\
		2. \url{https://doi.org/10.7910/DVN/NI3BDE} \\
		3. \url{https://doi.org/10.7910/DVN/D1N0GO} \\
		4. \url{https://doi.org/10.7910/DVN/IAOZOU}\\
\end{table}

% Note that some of this data has been aggregated in the Cumulative Contextual File, located here: \url{https://doi.org/10.7910/DVN/26451}.


Finally, other researchers have collected this data in forthcoming work. ``Separating Race and Party in Congressional
Elections'' by Bernard Fraga (available at \url{https://www.bernardfraga.com/research}) uses hand-coded race data from Congressional general and primary elections from 2006 - 2020. ``Partisanship and Nationalization in American Elections'' by Sharif Almani and Carlos Algara (\emph{Electoral Studies}, doi: \href{https://doi.org/10.7910/DVN/DGUMFI}{\texttt{10.7910/DVN/DGUMFI}}) contains county-level partisan election data from the statewide offices of President, Governor, and Senator going back to the Civil War.

\section{Reporting Errors and Suggesting Improvements}

If you find any errors or ambiguities, please do not hesitate to contact us. The preferred method is to open a ``Issue'' at \url{https://github.com/kuriwaki/cces_candidates/issues}, which we monitor. Otherwise, please email us at \href{mailto:jeremiahcha@g.harvard.edu}{\texttt{jeremiahcha@g.harvard.edu}}, \href{mailto:shirokuriwaki@gmail.com}{\texttt{shirokuriwaki@gmail.com}}, or \href{mailto:jsnyder@gov.harvard.edu}{\texttt{jsnyder@gov.harvard.edu}}.



\section{Version History}

\begin{itemize}
\item Dataverse v3.0: Added 2021, 2022, 2023 general elections (2024-10-18). These do not include House special elections held during this time.
\item Dataverse v2.0: Minor Fixes (2023-06-16). Fixes coding for fusion candidates, incumbents, special elections, and additional name inconsistencies (Github issues \href{https://github.com/kuriwaki/cces_candidates/issues/25}{25}, \href{https://github.com/kuriwaki/cces_candidates/issues/26}{26}, \href{https://github.com/kuriwaki/cces_candidates/issues/27}{27}, \href{https://github.com/kuriwaki/cces_candidates/issues/28}{28}, \href{https://github.com/kuriwaki/cces_candidates/issues/29}{29}, \href{https://github.com/kuriwaki/cces_candidates/issues/30}{30},
\href{https://github.com/kuriwaki/cces_candidates/issues/35}{35}). Thanks to Luca Bellodi and Bill Wilkerson for finding some of these errors.
\item Dataverse v1.0: Initial Dataverse Release (2021-09-27).
\end{itemize}

\section{Acknowledgements}

Our thanks to Stephen Ansolabehere and Jaclyn Kaslovsky for their suggestions and comments.



\end{document}

